Using Citation Data to Study Computer Science

### 2019-11-07 Scientometrics/comp.tex
m and methods

\title{An Appreciation of Computer Science%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
\subtitle{Article Level Clusters vs Journals\\ If so, write it here}

@@

\author{Sitaram Devarakonda         \and
        George Chacko %etc.
}

@@

\institute{Sitaram Devarakonda \at
              Netelabs, NET ESolutions Corporation, McLean, VA \\
              %Tel.: +123-45-678910\\
              %Fax: +123-45-678910\\
              \email{sitaram@nete.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           George Chacko \at
              Netelabs, NET ESolutions Corporation, McLean, VA \\
              \email{netelabs@nete.com}
}

@@

\begin{abstract}
Computer science and its applications has experienced dramatic growth and diversification over the last twenty years. Toward understanding the nature of these changes, we analyze a cohort of the computer science literature using a historiographic approach that relies on citations. To escape the imprecise definitions encoded in journal-level groupings, we build article-level clusters. Mention structure and information flow. We compare two different clustering approaches and reconcile them with subject categories in the Scopus database.

@@

Cite Alan Porter on expanding universe versus myopic focus. Cite Traag, Waltman, Lariviere papers. Peter Sjögårde and Per Ahlgren, Loet Leydesdorff, Caroline S. Wagner and Lutz Bornmann, Discontinuities in citation relations among journals: self-organized criticality as a model of scientific revolutions and change, Scientometrics, 10.1007/s11192-018-2734-6, 116, 1, (623-644), (2018). Antonio Perianes-Rodriguez and Javier Ruiz-Castillo, A comparison of the Web of Science and publication-level classification systems of science, Journal of Informetrics, 10.1016/j.joi.2016.10.007, 11, 1, (32-45), (2017). Qi Wang and Ludo Waltman, Large-scale analysis of the accuracy of the journal classification systems of Web of Science and Scopus, Journal of Informetrics, 10.1016/j.joi.2016.02.003, 10, 2, (347-364), (2016). Henry Small, Kevin W. Boyack and Richard Klavans, Identifying emerging topics in science and technology, 
Crossref
\section{Materials and Methods}
\label{sec:1}
Text with citations \cite{RefB} and \cite{RefJ}.

@@

\begin{itemize}
\item Spectral clustering
\begin{enumerate}
\item Subset Scopus by COMP
\item Match COMP to dblp at high stringency -> reduced number of scps
 \item Harvest cited refefences  from Scopus
 \item Cluster articles by direct citation using Graclus to a number that meets simple optimality criteria and does not impose cognitive challenges
 \item Reconcile cluster contents to journal classification for the purpose of referencing
 \item Infer inter-cluster relationships using direct citations, co-citations, and bibliographic coupling
\end{enumerate}
\item Co-citation based clustering
\begin{enumerate}
\item Stratify by year after subsetting by COMP and restricting to article and cp only
\item Apply fractional citation counting and identify highly-cited papers
\item Calculate normalized co-citations
\item Cluster by variable clustering method of Small
\item Evaluate major co-cited pairs
\end{enumerate}
\end{itemize}

>>>>

\title{Using Citation Data to Study Computer Science%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
\subtitle{Relationship With Other Fields}

@@

\author{Sitaram Devarakonda  \and
        Dmitriy Korobskiy \and
        Tandy Warnow \and
        George Chacko }

@@

\institute{Sitaram Devarakonda,  \at
              Netelabs, NET ESolutions Corporation, McLean, VA \\
              %Tel.: +123-45-678910\\
              %Fax: +123-45-678910\\
              \email{sitaram@nete.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
	  Dmitriy Korobskiy \at
              Netelabs, NET ESolutions Corporation, McLean, VA \\
              \email{dk@nete.com}
                  \and
           Tandy Warnow \at
              Dept of Computer Science, University of Iliinois Urbana-Champaign, Champaign IL \\
              \email{warnow@illinois.edu}
                  \and
           George Chacko \at
              Netelabs, NET ESolutions Corporation, McLean, VA \\
              \email{netelabs@nete.com}
}

@@

\begin{abstract}
Computer science and its applications has experienced dramatic growth and diversification over the last twenty years. Towards understanding the nature of these changes, we analyze a cohort of the computer science literature using a historiographic approach. To understand the features of this cohort and the flow of information within its components, we construct article level clusters whose elements are linked by citations or co-citations, and reconcile them with subject categories in the Scopus database.

@@

Cite Alan Porter on expanding universe versus myopic focus. Cite Traag, Waltman, Lariviere papers. Peter Sjögårde and Per Ahlgren, Loet Leydesdorff, Caroline S. Wagner and Lutz Bornmann, Discontinuities in citation relations among journals: self-organized criticality as a model of scientific revolutions and change, Scientometrics, 10.1007/s11192-018-2734-6, 116, 1, (623-644), (2018). Antonio Perianes-Rodriguez and Javier Ruiz-Castillo, A comparison of the Web of Science and publication-level classification systems of science, Journal of Informetrics, 10.1016/j.joi.2016.10.007, 11, 1, (32-45), (2017). Qi Wang and Ludo Waltman, Large-scale analysis of the accuracy of the journal classification systems of Web of Science and Scopus, Journal of Informetrics, 10.1016/j.joi.2016.02.003, 10, 2, (347-364), (2016). Henry Small, Kevin W. Boyack and Richard Klavans, Identifying emerging topics in science and technology, 
Crossref

%%%%%%%%%%%%%%%
\section{Materials and Methods}

\emph{Data} The dblp computer science bibliography, dblp-2018-08-01.xml.gz, consisting of 7,079,994 records was downloaded from  `https://dblp.uni-trier.de/xml/`. Publications were parsed from the xml source file and loaded into a PostgreSQL database. These parsed records were then matched to Scopus identifiers using PostgreSQL's full text search feature applied to document titles. The process involves parsing Scopus document titles into tokens, reducing the tokens into lexemes and returning a ts\_vector which lists the lexemes together with their positions. The ts\_vector is then matched with DBLP document titles. A ts\_rank is returned based on the frequency of matching lexemes. Only records in dblp with DOIs (4,291,130) were considered and only matches with a rank of $>$ 0.99 were accepted. This procedure resulted in a dataset of 634,100 publications. Cited references for these were extracted from Scopus, resulting in a total of 3,960,503 nodes (publications and references) and 11,992,387 edges that represents citations of these references and publications within the dataset. This dataset was labelled \emph{comp}.\par \emph{Clustering by Direct Citation} Graclus software was downloaded from \\`http://www.cs.utexas.edu/users/dml/Software/graclus.html'. The \emph{comp} dataset was cast as an undirected graph and stored in a file with a header line indicating the number of nodes and edges. Graclus accepts the number of clusters required as a parameter, so we used 10,15, 20, 25, 30, 40, and 50 as this parameter and  \emph{comp} as input. Based on criteria articulated in the Louvain and Leiden clustering algorithms, we chose 20 clusters as the number of output clusters. \par \emph{Clustering by Co-Citation} To build clusters using co-citations, we slightly modified the approach of Small and Sweeney (1985) as follows. We first calculated the number of citations accumulated across Scopus for all 634,100 articles in \emph{comp}. After discarding those publications that had not been cited at all, we restricted further analysis to the those in the 90th percentile or higher resulting in a dataset of size 50,869. We then identified all publications in Scopus that cite the 50,869 previously selected papers from \emph{comp}. The cited references in these papers were restricted to those in the set of 50,869 highly cited papers previously identified. For each publication in turn, all possible reference pairs were generated. The frequency of these co-cited pairs was then computed across the dataset  and normalized using Salton's cosine formula. A total of 2,208,135 unique co-cited pairs were thus obtained.

Variable level Clustering involves iteratively selecting the edges with highest normalized co-citation value and collecting
all the nodes related to them and so forth. One issue faced is the accumulation of one giant cluster.

To break them up we applied thresholds based on normalized co-citation values. In this experiment we started with 0.5 and progressively increased by 0.1
until 0.9, then by 0.01 until 0.99. Any cluster larger than 200 is carried over to be clustered based on higher thresholds.

After the completion of 1st round of clustering we had 6645 clusters with a total of 36024 nodes, losing 29.18 % nodes.
Clusters with only 2 nodes were eliminated bringing down the total clusters to 3760

To obtain target clusters of around 20 we did hierarchical clustering based on the above results. Inter-cluster relations were calculated for every
cluster and the max edge value based on normalized co-citaion is selected.

Based on max edge value two clusters which are merged into one. This process was applied for 9 times and we ended up with 12 clusters

First round input no of clusters: 3760
hierarchical\_clustering\_round,no\_of\_clusters
no1,1757
no2,853
no3,419
no4,208
no5,103
no6,51
no7,25
no8,12

Message Sitaram

\label{sec:1}
Text with citations \cite{RefB} and \cite{RefJ}.

@@

\begin{itemize}
\item Spectral clustering
\begin{enumerate}
\item Subset Scopus by COMP
\item Match COMP to dblp at high stringency -> reduced number of scps
 \item Harvest cited references  from Scopus
 \item Cluster articles by direct citation using Graclus to a number that meets simple optimality criteria and does not impose cognitive challenges
 \item Reconcile cluster contents to journal classification for the purpose of referencing
 \item Infer inter-cluster relationships using direct citations, co-citations, and bibliographic coupling
\end{enumerate}
\item Co-citation based clustering
\begin{enumerate}
\item Stratify by year after subsetting by COMP and restricting to article and cp only
\item Apply fractional citation counting and identify highly-cited papers
\item Calculate normalized co-citations
\item Cluster by variable clustering method of Small
\item Evaluate major co-cited pairs
\end{enumerate}
\end{itemize}
